---
title: "viz_part_01"
author: "Margaret"
date: "2022-10-10"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##lets import data

```{r}
library(tidyverse)
library(ggridges)
```

This is from the website

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

We are going to make a scatterplot now

```{r}
ggplot (weather_df, aes(x=tmin, y=tmax)) +
  geom_point()
```
Let make the same scatterplot, but different. The filtering step now indicates that we are only looking at central park data. 

```{r}
weather_df %>%
  drop_na()%>%
  filter (name =="CentralPark_NY")%>%
  ggplot (aes(x=tmin, y=tmax)) +
  geom_point()
```
Lets keep making the same plot but different. This now gives us the same blank plot as before. 

```{r}
weather_scatterplot =
  weather_df %>%
  drop_na()%>%
  ggplot (aes(x=tmin, y=tmax))
```

This then gives us the actual scatterplot now.
```{r}
weather_scatterplot + 
  geom_point()
```

#Lets make this fancier. This made it colorful. It colors it by the "name" variable

```{r}
weather_df%>%
  ggplot (aes(x=tmin, y=tmax, color = name)) +
  geom_point()
```

You can define aesthetic mapping for certain geometries. the geom smooth runs a line through the middle of your dataset to see if there is a linear association going on here in the data. 

```{r}
weather_df %>%
   ggplot (aes(x=tmin, y=tmax)) +
  geom_point(aes (color = name)) +
  geom_smooth()
  
```

This is the same plot but it makes the 3 different lines based on "name". it applied both to the scatterplot and the "name" variable. Anything in the ggplot() applis it to all scatterplots. If put in the geom_plot(), it only applies to the scatterplot 

```{r}
weather_df %>%
   ggplot (aes(x=tmin, y=tmax, color = name)) +
  geom_point() +
  geom_smooth()
```

It is hard to see everything going on in the plot. so in the geom_smooth(), we are going to define a specific parameter. We are not mapping alpha to a variable. We are defining alpha as .3 (alpha is between 0 and 1). Alpha is blending/making it more or less transparent with 0 being completely transparent.In this graph, the grey shaded regions on the lines are error bars. 

```{r}
weather_df %>%
   ggplot (aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth()
```

Getting rid of error bars.

```{r}
weather_df %>%
   ggplot (aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)
```
Maybe make separate panels. we are now going to tell ggplot make this plot but separate it by whatever variable we want to seaprate it by. This made it create 3 separate4 charts that is seaprated by the 3 names. In the facet_grid(), the left . defines the row and the name defines the columns of the charts. 
```{r}
weather_df %>%
   ggplot (aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name )
```

`tmax` and `tmin` is boring. Lets spice it up. Lets look at maximum trends of seasonal temp with the precipitation variabole be a part of this too. This shows the usual temp profile of these 3 diff places with precipitation. Now you have a precipiation variabole with diff colors. 

```{r}
weather_df%>%
  ggplot (aes(x = date, y = tmax, color = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)


```

Doing precipitation by color may not be totally useful. Here is it with size differences.

```{r}
weather_df%>%
  ggplot (aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```


## univariate plots
histograms, barplots, boxplots, violins....

This is not great bc it stacks the historgrams on top of each other. 
```{r}
weather_df %>%
  ggplot(aes( x = tmax, fill = name)) +
  geom_histogram()
```

Then we should separate the historgrams with facet_grid()
```{r}
weather_df %>%
  ggplot(aes( x = tmax, fill = name)) +
  geom_histogram() + 
  facet_grid(. ~ name)
```

Lets see more options. Instead of separating, use geom_density(). can also use ggplot(aes(x = tmax, fill = name)) to fill in the density plot by name. 
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, color = name)) +
  geom_density() 
```

Boxplots! 

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Violin plot- same as boxplot but also with density. This would be helpful if you are comparing 20-30 stations. 
```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```
OR you can take the density graphs and stack them on top of each other with geom_density_ridges(). 

```{r}
weather_df %>%
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```


##saving and embedding plots

```{r}
weather_scatterplot = 
  weather_df%>%
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth (se = FALSE) +
  facet_grid (. ~ name)

weather_scatterplot

ggsave("results/weather_scatterplot.pdf", weather_scatterplot, width = 8, height = 5)
```



